\bf Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-07-15 manuscript/main.tex
proof splits into propositions 1 and 2

	\begin{proposition}\label{proposition:eigen_noregul}
Let
$
(B^\ast, \Theta, \mathbf{y})
$
denote the observed data defined in \eqref{def:problem_instance}.
At the $l$-th iteration of the PLS algorithm, let
$( \hat{\mathbf{d}}_1^\top, \ldots, \hat{\mathbf{d}}_K^\top, \hat{\boldsymbol{\zeta}}^\top )^\top$
denote the coefficients of $\hat{\xi}$, defined in \eqref{def:maximizer_squared_empirical_cov},
which corresponds to the projected unit-norm direction that maximizes the squared empirical covariance.
These coefficients are obtained by solving the following generalized eigenvalue problem:
\begin{align}
	\left(
	\hat{\mathbf{d}}_1^\top, \ldots, \hat{\mathbf{d}}_K^\top, \hat{\boldsymbol{\zeta}}^\top
	\right)^\top
	=
	\arg \max_{\boldsymbol{\xi} \in \mathbb{R}^{MK+p}}   \boldsymbol{\xi}^\top V \boldsymbol{\xi}
	\quad \text{subject to} \quad \boldsymbol{\xi}^\top B^\ast \boldsymbol{\xi} = 1,
	\label{eq: Regularized generalized rayleigh quotient equation}
\end{align}
where
$$
V := \frac{1}{n^2} (B^\ast \Theta^\top \mathbf{y})(B^\ast \Theta^\top \mathbf{y})^\top \in \mathbb{R}^{(MK+p) \times (MK+p)},
$$

@@

	\section{Proofs of Propositions}
	$\boldsymbol{\cdot}$ \underline{Abbreviations}
	\begin{itemize}
		\item[--] ``CSI": Cauchy–Schwarz inequality
	\end{itemize}

	\subsection{Proof of Propositions \ref{proposition:eigen_noregul} and \ref{proposition:linear_noregul} }\label{section:proof:proposition:linear_noregul_system_unregularized}

@@

@@

	Therefore, problem of maximizing the squared covariance can be written as
			\widehat{\textnormal{Cov}}^2 (
			\bigl \langle
			(\Theta_1^\top \mathbf{y} , \ldots, \Theta_K^\top \mathbf{y}, Z^\top \mathbf{y}),
			(\mathbf{d}_1, \ldots, \mathbf{d}_K, \boldsymbol{\zeta})
			\bigr\rangle_B^2.
	%Therefore, the optimization problem 
	presented in Proposition \ref{proposition:linear_noregul_system_unregularized} is equivalent to the following constrained optimization problem on the Euclidean space:

@@

@@

	Let us denote the objective function as
	f(\{\mathbf{d}_j\}, \boldsymbol{\zeta}) := 

@@

	\left(
	\sum_{j=1}^K \mathbf{y}^\top \Theta_j B \mathbf{d}_j + \mathbf{y}^\top Z \boldsymbol{\zeta}
	\right)^2,
	and the constraint function as
	g(\{\mathbf{d}_j\}, \boldsymbol{\zeta}) := \sum_{j=1}^K \mathbf{d}_j^\top B \mathbf{d}_j + \boldsymbol{\zeta}^\top \boldsymbol{\zeta} - 1 = 0.

	The feasible set is compact, as it corresponds to the boundary of an ellipsoid in a finite-dimensional Euclidean space. Since the objective function $f$ is continuous and the constraint defines a compact feasible set, the  Weierstrass Extreme Value Theorem (see, for example, Theorem 2.3.1 in \citealt{bazaraaNonlinearProgrammingTheory2006}) ensures the existence of a global maximizer.

	Define the Lagrangian:
	\[
	\mathcal{L}(\{\mathbf{d}_j\}, \boldsymbol{\zeta}, \lambda) 
	:= 
	f(\{\mathbf{d}_j\}, \boldsymbol{\zeta}) - \lambda g(\{\mathbf{d}_j\}, \boldsymbol{\zeta}).
	\]
	The constraint function $g$ is continuously differentiable, and its gradient vanishes only at the origin. Since the origin is not feasible (the constraint requires norm 1), the gradient of $g$ is nonzero at all feasible points. Thus, the linear independence constraint qualification (LICQ) holds, ensuring that the KKT conditions are necessary for local maxima (see, for example, Theorem 5.3.1 in \citealt{bazaraaNonlinearProgrammingTheory2006}).
	Taking gradients of $\mathcal{L}$ and setting them to zero yields:
	\begin{align*}
		\nabla_{\mathbf{d}_j} \mathcal{L} &= 
		\frac{2}{n^2} 
		\left(
		\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \mathbf{d}_\ell + \mathbf{y}^\top Z \boldsymbol{\zeta}
		\right)
		B \Theta_j^\top \mathbf{y}
		- 2\lambda B \mathbf{d}_j = 0, \quad \forall j, 
		\numberthis \label{lag_grad_d}
		\\
		\nabla_{\boldsymbol{\zeta}} \mathcal{L} &= 
		\frac{2}{n^2} 
		\left(
		\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \mathbf{d}_\ell + \mathbf{y}^\top Z \boldsymbol{\zeta}
		\right)
		Z^\top \mathbf{y}
		- 2\lambda \boldsymbol{\zeta} = 0.
		\numberthis \label{lag_grad_zeta}
	\end{align*}
	By pre-multiplying equation \eqref{lag_grad_d} by $B^{-1}$ and observing that the scalar quantity $n^{-2}\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \mathbf{d}_\ell + \mathbf{y}^\top Z \boldsymbol{\zeta}$ appears in both \eqref{lag_grad_d} and \eqref{lag_grad_zeta}, the gradient conditions together imply that any local optimum must be  scalar multiples of
	$
	\Theta_1^\top \mathbf{y}, \ldots, \Theta_K^\top \mathbf{y}, Z^\top \mathbf{y},
	$
	all scaled by the same factor. There exists only one set of vectors  that also satisfies the primal equality constraint, namely the normalized version of the vector set above. The global maximizer is therefore given by
	\begin{equation*}
		\mathbf{d}_j^\ast = \frac{1}{\sqrt{\sum_{j=1}^K 
				\mathbf{y}^\top
				\Theta_j
				B \Theta_j^\top \mathbf{y} 
				+ 
				\mathbf{y}^\top Z Z^\top \mathbf{y}}}
		\Theta_j^\top \mathbf{y},
		\quad
		%%%%%%%%%%%%%%%%
		\boldsymbol{\zeta}^\ast = \frac{1}{\sqrt{\sum_{j=1}^K 
				\mathbf{y}^\top
				\Theta_j
				B \Theta_j^\top \mathbf{y} 
				+ 
				\mathbf{y}^\top Z Z^\top \mathbf{y}}}
		Z^\top \mathbf{y},
	\end{equation*}
	which are the same as  $\hat{\mathbf{d}}_1, \ldots, \hat{\mathbf{d}}_K$ and $\hat{\boldsymbol{\zeta}}$. This completes the proof of  Proposition \ref{proposition:linear_noregul_system_unregularized}.

>>>>

	\begin{proposition}\label{proposition:eigen_noregul}
Let
$
(B^\ast, \Theta, \mathbf{y})
$
denote the observed data defined in \eqref{def:problem_instance}.
At the $l$-th iteration of the PLS algorithm, let
$( \hat{\mathbf{d}}_1^\top, \ldots, \hat{\mathbf{d}}_K^\top, \hat{\boldsymbol{\zeta}}^\top )^\top$
denote the coefficients of $\hat{\xi}$, defined in \eqref{def:maximizer_squared_empirical_cov},
which corresponds to the projected unit-norm direction that maximizes the squared empirical covariance.
These coefficients are obtained by solving the following generalized eigenvalue problem:
\begin{equation}	\label{eq: Regularized generalized rayleigh quotient equation}
	\left(
	\hat{\mathbf{d}}_1^\top, \ldots, \hat{\mathbf{d}}_K^\top, \hat{\boldsymbol{\zeta}}^\top
	\right)^\top
	=
	\arg \max_{\boldsymbol{\xi} \in \mathbb{R}^{MK+p}}   \boldsymbol{\xi}^\top V \boldsymbol{\xi}
	\quad \text{subject to} \quad \boldsymbol{\xi}^\top B^\ast \boldsymbol{\xi} = 1,
\end{equation}
where
\begin{equation}\label{def:V_matrix}
V := \frac{1}{n^2} (B^\ast \Theta^\top \mathbf{y})(B^\ast \Theta^\top \mathbf{y})^\top \in \mathbb{R}^{(MK+p) \times (MK+p)},
\end{equation}

@@

	\section{Proof of Proposition \ref{proposition:eigen_noregul} }\label{section:proof:proposition:linear_noregul_system_unregularized}

@@

	\begin{proof}

@@

	Therefore, problem of maximizing the squared covariance  is equivalent to the following constrained optimization problem on the Euclidean space:

@@

		\numberthis \label{optim:summation_form}

@@

The objective function matches that of \eqref{eq: Regularized generalized rayleigh quotient equation}, taking the form of a quadratic expression involving the matrix $V$ defined in \eqref{def:V_matrix}:

@@

\biggr)^2
%%%%%%%%%%%%%%%%%%%%
\Theta
B^\ast  
\boldsymbol{\xi}
)^2
\\
&	= 
\boldsymbol{\xi}^\top
(B^\ast \Theta^\top \mathbf{y})(B^\ast \Theta^\top \mathbf{y})^\top 
\boldsymbol{\xi}
	\boldsymbol{\xi}^\top
V 
\boldsymbol{\xi},
and the constraint is identical to that in \eqref{eq: Regularized generalized rayleigh quotient equation}, because we have:
	+ \boldsymbol{\zeta}^\top \boldsymbol{\zeta}
	\boldsymbol{\xi}^\top B^\ast \boldsymbol{\xi}.
Therefore, the optimization problem \eqref{optim:summation_form}, which is equivalent to the squared covariance maximization problem, reduces to the generalized eigenvalue problem \eqref{eq: Regularized generalized rayleigh quotient equation}. This completes the proof of Proposition~\ref{proposition:eigen_noregul}.
 \end{proof}

 \section{Proof of Proposition  \ref{proposition:linear_noregul}}
 \begin{proof}
	From \eqref{optim:summation_form}, let us denote the objective function as
	\[
	f(\{\mathbf{d}_j\}, \boldsymbol{\zeta}) := 
	\frac{1}{n^2}
	\left(
	\sum_{j=1}^K \mathbf{y}^\top \Theta_j B \mathbf{d}_j + \mathbf{y}^\top Z \boldsymbol{\zeta}
	\right)^2,
	\]
	and the constraint function as
	\[
	g(\{\mathbf{d}_j\}, \boldsymbol{\zeta}) := \sum_{j=1}^K \mathbf{d}_j^\top B \mathbf{d}_j + \boldsymbol{\zeta}^\top \boldsymbol{\zeta} - 1 = 0.
	\]
	The feasible set is compact, as it corresponds to the boundary of an ellipsoid in a finite-dimensional Euclidean space. Since the objective function $f$ is continuous and the constraint defines a compact feasible set, the  Weierstrass Extreme Value Theorem (see, for example, Theorem 2.3.1 in \citealt{bazaraaNonlinearProgrammingTheory2006}) ensures the existence of a global maximizer.

	Define the Lagrangian:
	\[
	\mathcal{L}(\{\mathbf{d}_j\}, \boldsymbol{\zeta}, \lambda) 
	:= 
	f(\{\mathbf{d}_j\}, \boldsymbol{\zeta}) - \lambda g(\{\mathbf{d}_j\}, \boldsymbol{\zeta}).
	\]
	The constraint function $g$ is continuously differentiable, and its gradient vanishes only at the origin. Since the origin is not feasible (the constraint requires norm 1), the gradient of $g$ is nonzero at all feasible points. Thus, the linear independence constraint qualification (LICQ) holds, ensuring that the KKT conditions are necessary for local maxima (see, for example, Theorem 5.3.1 in \citealt{bazaraaNonlinearProgrammingTheory2006}).
	Taking gradients of $\mathcal{L}$ and setting them to zero yields:
	\begin{align*}
		\nabla_{\mathbf{d}_j} \mathcal{L} &= 
		\frac{2}{n^2} 
		\left(
		\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \mathbf{d}_\ell + \mathbf{y}^\top Z \boldsymbol{\zeta}
		\right)
		B \Theta_j^\top \mathbf{y}
		- 2\lambda B \mathbf{d}_j = 0, \quad \forall j, 
		\numberthis \label{lag_grad_d}
		\\
		\nabla_{\boldsymbol{\zeta}} \mathcal{L} &= 
		\frac{2}{n^2} 
		\left(
		\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \mathbf{d}_\ell + \mathbf{y}^\top Z \boldsymbol{\zeta}
		\right)
		Z^\top \mathbf{y}
		- 2\lambda \boldsymbol{\zeta} = 0.
		\numberthis \label{lag_grad_zeta}
	\end{align*}
	By pre-multiplying equation \eqref{lag_grad_d} by $B^{-1}$ and observing that the scalar quantity $n^{-2}\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \mathbf{d}_\ell + \mathbf{y}^\top Z \boldsymbol{\zeta}$ appears in both \eqref{lag_grad_d} and \eqref{lag_grad_zeta}, the gradient conditions together imply that any local optimum must be  scalar multiples of
	$
	\Theta_1^\top \mathbf{y}, \ldots, \Theta_K^\top \mathbf{y}, Z^\top \mathbf{y},
	$
	all scaled by the same factor. There exists only one set of vectors  that also satisfies the primal equality constraint, namely the normalized version of the vector set above. The global maximizer is therefore given by
	\begin{equation*}
		\mathbf{d}_j^\ast = \frac{1}{\sqrt{\sum_{j=1}^K 
				\mathbf{y}^\top
				\Theta_j
				B \Theta_j^\top \mathbf{y} 
				+ 
				\mathbf{y}^\top Z Z^\top \mathbf{y}}}
		\Theta_j^\top \mathbf{y},
		\quad
		%%%%%%%%%%%%%%%%
		\boldsymbol{\zeta}^\ast = \frac{1}{\sqrt{\sum_{j=1}^K 
				\mathbf{y}^\top
				\Theta_j
				B \Theta_j^\top \mathbf{y} 
				+ 
				\mathbf{y}^\top Z Z^\top \mathbf{y}}}
		Z^\top \mathbf{y},
	\end{equation*}
	which are the same as  $\hat{\mathbf{d}}_1, \ldots, \hat{\mathbf{d}}_K$ and $\hat{\boldsymbol{\zeta}}$. This completes the proof of  Proposition \ref{proposition:linear_noregul}.
	\end{proof}
